# Reproducible Research: Peer Assessment 1

```{r}
library("Hmisc")

## Loading and preprocessing the data
```{r}
unzip(zipfile="activity.zip")
setwd("/home/chase/repro/RepData_PeerAssessment1/")
activitydata <- read.csv("activity.csv",header=TRUE)
activityday <- transform(activitydata,date=as.Date(date))
time <- formatC(activitydata$interval/100, 2, format = "f")
activitydata$date.time <- as.POSIXct(paste(activitydata$date, time), format = "%Y-%m-%d %H.%M", 
    tz = "GMT")
activitydata$time <- format(activitydata$date.time, format = "%H:%M:%S")
activitydata$time <- as.POSIXct(activitydata$time, format = "%H:%M:%S")
```

## What is mean total number of steps taken per day?
First, Lets transform the data a bit to give us totals by day.

```{r}
steps <- tapply(activityday$steps, activityday$date, FUN = sum,na.rm=TRUE)
```
Now, Let's Plot This to a Histogram.

```{r fig.width=7, fig.height=6}
hist(steps,breaks=61,main="Steps Per Day By Frequency",xlab="Daily Steps")
```

Now, Let's calculate Mean and Median using these day totals.

```{r}
mean(steps,na.rm=TRUE)
median(steps,na.rm=TRUE)
```

## What is the average daily activity pattern?

First, we are going to calculate the averages for each interval.
```{r}
interval <- tapply(activitydata$steps, activitydata$time,mean,na.rm=TRUE)
```

Now we load it into a data frame.

```{r}
patterndata <- data.frame(time = as.POSIXct(names(interval)), interval = interval)
```

```{r fig.width=7, fig.height=6}
plot(y=patterndata$interval,x=patterndata$time,type='l',xlab="Time of Day",ylab="Mean Number of Steps")
```

Now Lets Find the five minute interval with the highest number of steps.

```{r}
which.max(patterndata$interval)
```

## Imputing missing values

Let's find all the missing values.

```{r}
sum(is.na(activitydata$steps))
```

Now let's fill in the missing data with the average for that interval.

```{r}
filledactivity <- activitydata
filledactivity$steps <- with(filledactivity,impute(steps,mean))
filledactivityday <- transform(filledactivity,date=as.Date(date))
filledsteps <- tapply(filledactivityday$steps, filledactivityday$date, FUN = sum,na.rm=TRUE)
```
Now, Let's Plot This to a Histogram.

```{r fig.width=7, fig.height=6}
hist(filledsteps,breaks=61,main="Steps Per Day By Frequency",xlab="Daily Steps")
```

Now, Let's calculate Mean and Median using these day totals.

```{r}
mean(filledsteps)
median(filledsteps)
```

As we can see, imputing the NA values has raised the mean and median.


## Are there differences in activity patterns between weekdays and weekends?

Let's isolate the weekends from the weekdays.

```{r}
weekday <- function(date) {
    if (weekdays(date) %in% c("Saturday", "Sunday")) {
        return("weekend")
    } else {
        return("weekday")
    }
}

filledactivity$weekday <- as.factor(sapply(filledactivity$date.time, weekday))

```

Let's subset the data into weekend and weekday datasets

```{r}
weekends <- filledactivity[filledactivity$weekday=="weekend",]
weekday <- filledactivity[filledactivity$weekday=="weekday",]

weekday <- as.vector(tapply(weekday$steps, weekday$interval, mean))
weekends <- as.vector(tapply(weekends$steps, weekends$interval, mean))

```

Now Let's Plot it

```{r}
int <- as.numeric(filledactivity$interval[1:288])

par(mfrow=c(2,1))
plot(int,weekends,type='l',main="Weekend",ylab="Steps",xlab="Interval")
plot(int,weekday,type='l',main="Weekday",ylab="Steps",xlab="Interval")
